A More Complete Algebra for File Synchronization

### 2015-11-18 2015/paper.tex
Updated proof to new notation

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}}
\newcommand{\setd}{\setvx{D}}
\newcommand{\setb}{\setvx{\empt}}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\toppath}{\mathtt{None}} % return value of parent()
\newcommand{\fsbroken}{\mathtt{Broken}} % broken FS value
\newcommand{\FS}{\Phi} % a filesystem
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\pp}{p\!\!\uparrow} % parent path

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

@@

@@

\begin{myrul}
\[\infer{A\equiv B}{A\eqext B}\]
\end{myrul}

@@

\begin{myrul}
\[\infer{A\eqnrw B}{\worksc{A}{B}}\]
\end{myrul}

\begin{myrul}
\[\infer{\works{A_1,\ldots,A_{n-1},A_n}}{\works{A_1,\ldots,A_{n-1}}}\]
\end{myrul}

@@

@@

Pairs of commands in general will have the form
\[ [\cxy(p,V_Y); \czw(q,V_W)] \]
where $X,Y,Z,W\in\typeset$, $p,q\in\setp$, 
and values are of the appropriate type: 
$V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$.

The following axioms describe multiple statements
that we accept as true in the algebra.
These statements are generated by substituting 
all possible values for
$X,Y,Z,W\in\typeset$; $p,q\in\setp$; $V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$
under the conditions noted in the axiom.

\begin{myax}\label{ax_separate_commute}
$\cxy(p,V_Y); \czw(q,V_W) \equiv \czw(q,V_W); \cxy(p,V_Y)$ where $p\not\descendant q$ and $q\not\descendant p$.
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
$\cxy(p,V_Y); \czw(q,V_W) \nequiv \cbrk$ where $p\not\descendant q$ and $q\not\descendant p$.
\end{myax}

\begin{myax}\label{ax_same_breaks}
$\cxy(p,V_Y); \czw(p,V_W) \equiv \cbrk$ if $Y\ne Z$.
\end{myax}

\begin{myax}\label{ax_distantchild_breaks}
$\cxy(p,V_Y); \czw(q,V_W) \equiv \cbrk$ if $p\descendant q$ and $p\neq\parent(q)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$.
\end{myax}

\begin{myax}\label{ax_distantparent_breaks}
$\cxy(p,V_Y); \czw(q,V_W) \equiv \cbrk$ if $p\ancestor q$ and $\parent(p)\neq q$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$.
\end{myax}

\begin{mydef}{Construction pair.}
A pair of commands on paths $p$ and $q$ is a construction pair if $p=\parent(q)$ and the commands are one
of the following:
   \begin{gather*}
   % bd>bf
            [\cbd(p, D); \cbf(q, F)] \\
   % bd>bd
            [\cbd(p, D_1); \cbd(q, D_2)] \\
   % fd>bf
            [\cfd(p, D); \cbf(q, F)] \\
   % fd>bd
            [\cfd(p, D_1); \cbd(q, D_2)]
   \end{gather*}
\end{mydef}

\begin{mydef}{Destruction pair.}
A pair of commands on paths $p$ and $q$ is a destruction pair if $\parent(p)=q$ and the commands are one
of the following:
   \begin{gather*}
   % fb<db ??
            [\cfb(p); \cdb(\pp)] \\
   % fb<df ??
            [\cfb(p); \cdf(\pp, F)] \\
   % db<db ??
            [\cdb(p); \cdb(\pp)] \\
   % db<df ??
            [\cdb(p); \cdf(\pp, F)]
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directchild_breaks}
$\cxy(p,V_Y); \czw(q,V_W) \equiv \cbrk$ if $p=\parent(q)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
\end{myax}

\begin{myax}\label{ax_directparent_breaks}
$\cxy(p,V_Y); \czw(q,V_W) \equiv \cbrk$ if $\parent(p)=q$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
\end{myax}

@@

% Separate:
% * commute
% - if both are BB or DD, extended by empty seq
% - if either is BB or DD, extended by the other command
% - not equiv to break
% Same:
% * equiv break unless 1-out == 2-in
% - extended by empty seq if 1-out==2-in & 1-in==2-out==B,D
% - extended by single command (if 1-out==2-in) (otherwise breaks so true)

@@

\section{Extending the algebra}

We note the following axioms:

\begin{itemize}
\item Laws for construction assertion:
   \begin{gather*}
            [\cbf(p, F)] \equiv [\cdd(\pp); \cbf(p, F)]\\
            [\cbd(p, D)] \equiv [\cdd(\pp); \cbd(p, D)]
   \end{gather*}
\item Laws for destruction assertion:
   \begin{gather*}
            [\cdb(\pp)] \equiv [\cbb(p); \cdb(\pp)]\\
            [\cdf(\pp, F)] \equiv [\cbb(p); \cdf(\pp, F)]
   \end{gather*}
\end{itemize}

Separately, we extend the algebra with the following judgements and axioms:

\begin{mydef}{$A\coworks B$}
or the domains of $A$ and $B$ overlap.
Its intended interpretation is that
there is at least one filesystem neither $A$ nor $B$ breaks.
It is a commutative (but not transitive) relation.
For $A\ncoworks B$ we say that $A$ and $B$ are disjoint.
\end{mydef}

\begin{mydef}{$A\wrext B$}
or the domain of $B$ extends that of $A$.
Its intended interpretation is that
$B$ does not break any filesystem $A$ does not break.
It is a transitive relation.
\end{mydef}

% TODO See above
\begin{myax}$A\coworks B \Rightarrow A\nequiv\cbrk$\end{myax}
\begin{myax}$A\eqext B \Rightarrow A\wrext B$\end{myax}
\begin{myax}\label{axiom:wrextpostfix}$[A;S] \wrext A$\end{myax}
\begin{myax}$A\coworks B\wrext C \Rightarrow A\coworks C$\end{myax}
\begin{myax}\label{axiom:ncowwrnrw}$A\ncoworks B\wrnrw C \Rightarrow A\ncoworks C$\end{myax}

\begin{mylem}\label{lemma:cowdelfix}$[A;S]\coworks B \Rightarrow A\coworks B$\end{mylem}
\begin{mylem}\label{lemma:ncowpostfix}$A\ncoworks B \Rightarrow [A;S]\ncoworks B$\end{mylem}
\begin{proof}From \cref{axiom:wrextpostfix,axiom:ncowwrnrw}\end{proof}

\begin{mylem}\label{lemma:cowinputmatch}
If $A$ and $B$ are minimal sequences, $A\coworks B$,
and there are commands on path $p$ in both $A$ ($\cxy(p, V_Y)\in A$) and $B$ ($\czw(p, V_W)\in B$)
then their input types must match ($X=Z$).
TODO: Can this be derived from the algebra?
TODO: maybe this is the only thing needed
\end{mylem}

@@

\begin{proof}
First we prove the following.
Mark all commands in $A$ that also appear in $B$.
Then $\exists A^* \in \orderset{A}$ in which the marked commands are at the beginning.

@@

We show $A$ can be transformed via equivalences so that this would be true
by induction on the number of non-marked commands preceding the last marked command.
If it is 0, no transformation is needed.
For the induction step, select the last marked command preceded by a non-marked command.
If the two commute, reverse them, thereby reducing the number on relevant non-marked commands.
If they do not commute, they must be a construction or destruction pair as we know
that $A$ is minimal and $A\nequiv\cbrk$.
Therefore either $q=\parent(p)$ or $p=\parent(q)$.

@@

If $B$ has a command on $q$, then
from \cref{lemma:neighbour}
we know that there is a $B'$ in $\orderset{B}$ where it is next to $\cxy$,
and from \cref{lemma:minextset,lemma:minextorder} we know that $B'\equiv B$.
From the construction and destruction laws we also know that a command on $q$
must precede $\cxy(p, V_Y)$ just as it does so in $A$.

If $B$ has no command on $q$, then let $B'$ be $B$, but with an extra command added to it just before $\cxy$
according to the construction or destruction assertion laws, from which we know
that $B'\equiv B$.
We also know that $B'$ is still minimal.

We therefore have a minimal $B'$ which has a command on $q$ just before $\cxy(p, V_Y)$ and $B'\equiv B$.
Let this command on $q$ be $\cqr(q, V_R)$.
\begin{gather*}
A = [\cdots; \czw(q, V_W); \cxy(p, V_Y); \cdots] \\
B' = [\cdots; \cqr(q, V_R); \cxy(p, V_Y); \cdots]
\end{gather*}

As $A\coworks B\equiv B'$, 
from \cref{lemma:cowinputmatch}
we know that $Q=Z$.

Going back to the construction and destruction laws we see that the output type of the first command
is always the same, and it is either $D$ or $\empt$. Therefore $R=W$ and $\cqr(q, V_R)=\czw(q, V_W)$
as $|\setd|=|\setb|=1$. 
If $B$ originally had a command on $q$,
this is a contradiction as $\czw$ was not marked, but we see it must also be in $B'$ and therefore in $B$.
If $B$ had no command on $q$,
this is a contradiction because the command added to $B$ is an assertion command, so $Z=Q=R=W$, 
but $A$ contained no assertion commands.

@@

We now know that there is an $A'\equiv A$ in which commands in $A\cap B$
are at the beginning, and therefore 
from \cref{axiom:wrextpostfix}
$A\wrext \ordered{A\cap B}$ and by symmetry $B\wrext \ordered{A\cap B}$.
\end{proof}

>>>>

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}}
\newcommand{\setd}{\setvx{D}}
\newcommand{\setb}{\setvx{\empt}}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\toppath}{\mathtt{None}} % return value of parent()
\newcommand{\fsbroken}{\mathtt{Broken}} % broken FS value
\newcommand{\FS}{\Phi} % a filesystem
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\pp}{p^*} % parent path

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\cxw}{\fscommand{X}{W}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

@@

\begin{mydef}{Unrelated paths.}
Paths $p$ and $q$ are unrelated if $p\not\descendant q$ and $p\not\ancestor q$.
\end{mydef}

@@

\begin{myrul}\label{r_eq_ex}
\[\infer{A\equiv B}{A\eqext B}\]
\end{myrul}

@@

\begin{myrul}\label{r_ex_w}
\[\infer{A\eqnrw B}{\worksc{A}{B}}\]
\end{myrul}

These rules also apply if the second, conditional part of $\worksc{\ldots}{B_1,\ldots,B_n}$ is empty:

@@

\begin{myrul}
\[\infer{\worksc{A_1,\ldots,A_i,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{A_i,A_1,\ldots,A_{i-1},A_{i+1},\ldots,A_n}{B_1,\ldots,B_n}}\]
\end{myrul}

\begin{myrul}\label{r_w_w}
\[\inferr{\worksc{A_1,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{C}{A_n}}{\worksc{A_1,\ldots,A_n,C}{B_1,\ldots,B_n}}\]
\end{myrul}

@@

Pairs of commands in general will have the form
\[ \cxy(p,V_Y); \czw(q,V_W) \]
where $X,Y,Z,W\in\typeset$, $p,q\in\setp$, 
and values are of the appropriate type: 
$V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$.

The following axioms describe multiple statements
that we accept as true in the algebra.
These statements are generated by substituting 
all possible values for
$X,Y,Z,W\in\typeset$; $p,q,\pp\in\setp$; $V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$
under the conditions noted in the axiom.

\begin{myax}\label{ax_separate_commute}
Commands on unrelated paths commute:
$\cxy(p,V_Y); \czw(q,V_W) \equiv \czw(q,V_W); \cxy(p,V_Y)$ where $p\not\descendant q$ and $q\not\descendant p$.
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated paths do not break every filesystem:
$\cxy(p,V_Y); \czw(q,V_W) \nequiv \cbrk$ where $p\not\descendant q$ and $q\not\descendant p$.
\end{myax}

\begin{myax}\label{ax_same_breaks}
Commands on the same path break every filesystem if incompatible:
$\cxy(p,V_Y); \czw(p,V_W) \equiv \cbrk$ where $Y\ne Z$.
\end{myax}

\begin{myax}\label{ax_same_emptyseq}
Commands on the same path simplify:
$\cxy(p,V_Y); \czw(p,V_W) \eqext \emptyseq$ where $Y=Z$ and $X=W=\fsregcommandchar{\empt}$ or $X=W=\fsregcommandchar{D}$.
\end{myax}

\begin{myax}\label{ax_same_singlec}
Commands on the same path simplify:
$\cxy(p,V_Y); \czw(p,V_W) \equiv \cxw(p,V_W)$ where $Y=Z$ and $XW\neq\cbb$ and $XW\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_distantchild_breaks}
$\cxy(\pp,V_Y); \czw(p,V_W) \equiv \cbrk$ where $\pp\descendant p$ and $\pp\neq\parent(p)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$.
\end{myax}

\begin{myax}\label{ax_distantparent_breaks}
$\cxy(p,V_Y); \czw(\pp,V_W) \equiv \cbrk$ where $p\ancestor \pp$ and $\parent(p)\neq \pp$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$.
\end{myax}

\begin{mydef}{Construction pair.}
A pair of commands on paths $\pp$ and $p$ is a construction pair if $\pp=\parent(p)$ and the commands are one
of the following:
   \begin{gather*}
   % bd>bf
            [\cbd(\pp, D); \cbf(p, F)] \\
   % bd>bd
            [\cbd(\pp, D_1); \cbd(p, D_2)] \\
   % fd>bf
            [\cfd(\pp, D); \cbf(p, F)] \\
   % fd>bd
            [\cfd(\pp, D_1); \cbd(p, D_2)]
   \end{gather*}
\end{mydef}

\begin{mydef}{Destruction pair.}
A pair of commands on paths $p$ and $\pp$ is a destruction pair if $\parent(p)=\pp$ and the commands are one
of the following:
   \begin{gather*}
   % fb<db ??
            [\cfb(p); \cdb(\pp)] \\
   % fb<df ??
            [\cfb(p); \cdf(\pp, F)] \\
   % db<db ??
            [\cdb(p); \cdb(\pp)] \\
   % db<df ??
            [\cdb(p); \cdf(\pp, F)]
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directchild_breaks}
$\cxy(\pp,V_Y); \czw(p,V_W) \equiv \cbrk$ where $\pp=\parent(p)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
\end{myax}

\begin{myax}\label{ax_directparent_breaks}
$\cxy(p,V_Y); \czw(\pp,V_W) \equiv \cbrk$ where $\parent(p)=\pp$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
\end{myax}

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a child path:
$\cbb(p);\cxy(\pp,V_Y) \equiv \cxy(\pp,V_Y) \equiv \cxy(\pp,V_Y);\cbb(p)$ where $\pp\descendant p$ and $\cxy\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_parent_assert}
An assertion command can be added on the parent path:
$\cdd(\pp);\cxy(p,V_Y) \equiv \cxy(p,V_Y) \equiv \cxy(p,V_Y);\cdd(\pp)$ where $\pp\descendant p$ and $\cxy\neq\cbb$.
\end{myax}

\begin{myax}\label{ax_assert}
$\cxy(p,V_Y) \eqext \emptyseq$ where $\cxy=\cbb$ or $\cxy=\cdd$.
\end{myax}

@@

@@

@@

\begin{proof}
First we prove the following.
Mark all commands in $A$ that also appear in $B$.
Then there is an $A'$ in $\orderset{A}$ in which the marked commands are at the beginning.

@@

We show $A$ can be transformed via equivalences so that this would be true,
by induction on the number of non-marked commands preceding the last marked command.
If it is 0, no transformation is needed.
For the induction step, consider the last marked command preceded by a non-marked command,
and the non-marked command itself.
As $A$ is a simple command and $\works{A}$, from 
\cref{ax_distantchild_breaks,ax_distantparent_breaks,ax_directchild_breaks,ax_directparent_breaks}
we know that these commands can only be on unrelated paths or form a construction or destruction pair.
In the first case, swapping the commands results in a sequence equivalent to $A$,
with the number on relevant non-marked commands reduced by one.
In the last two cases,
either $q=\parent(p)$ or $p=\parent(q)$.

@@

If $B$ has a command on $q$, then
from \cref{lemma:neighbour}
we know that there is a $B'$ in $\orderset{B}$ where it is next to $\cxy$,
and from \cref{lemma:minextset,lemma:minextorder} we know that $B'\equiv B$.
% TODO Use when describing ordering:
Considering the eight construction and destruction pairs we see that the command
on the parent path determines which group the pair belongs to.
This in turn determines whether the command on the child path must
precede or follow the other command as otherwise the sequence would break all filesystems.
This argument holds for both $A$ and $B$, and so the command on $q$ must be on
the same side of the command on $p$ in both sequences.

If $B$ has no command on $q$, then let $B'$ be $B$, but with an extra command added to it just before $\cxy$
according to \cref{ax_child_assert,ax_parent_assert}, from which we know
that $B'\equiv B$.
We also know that $B'$ is still minimal (but no longer simple).

We therefore have a minimal $B'$ which has a command on $q$ just before $\cxy(p, V_Y)$ and $B'\equiv B$.
Let this command on $q$ be $\cqr(q, V_R)$.
\begin{gather*}
A = \cdots; \czw(q, V_W); \cxy(p, V_Y); \cdots \\
B' = \cdots; \cqr(q, V_R); \cxy(p, V_Y); \cdots
\end{gather*}

As $\works{A,B}$ and so $\works{A,B'}$ (via \cref{r_eq_ex,r_ex_w,r_w_w}), 
from \cref{worksinputmatch}
we know that $Q=Z$.

Going back to the construction and destruction pairs we see that the output type of the first command
is always the same, and it is either $D$ or $\empt$ depending on the relationship between $p$ and $q$.
Therefore $R=W$ and $\cqr(q, V_R)=\czw(q, V_W)$
as $|\setd|=|\setb|=1$. 
If $B$ originally had a command on $q$,
this is a contradiction as $\czw$ was not marked, but we see it must also be in $B'$ and therefore in $B$.
If $B$ had no command on $q$,
this is a contradiction because the command on $q$ in $B'$ is an assertion command, so $Z=Q=R=W$, 
but $A$ contained no assertion commands.

@@

We now know that there is an $A'\equiv A$ in which commands in $A\cap B$
are at the beginning, and therefore 
from \cref{worksextpostfix}
$\worksc{\ordered{A\cap B}}{A}$ and by symmetry $\worksc{\ordered{A\cap B}}{B}$.
\end{proof}
